Creating the ``r params$package_name`` R package

### 2025-06-26 create-FSHybridPLS.Rmd
started manuscript logging

@@

## asd
### get_pre_corrcov
construct the matrix V_star:
$$
    V^\ast = n^{-2}\begin{bmatrix}
         J \widetilde{C}^\top \mathbf{y}\mathbf{y}^\top \widetilde{C} J & J \widetilde{C}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z} \\
        \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C} J &   \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z}
    \end{bmatrix} = n^{-2} J^* \widetilde{C}^{*\top} \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C}^{*} J^*
    \in \mathbb{R}^{(MK+p) \times (MK+p)},
$$
where $J$ is defined in `get_gram_matrix_block`
equation (7), 
```{r}
#' Compute V matrix (pre-corrected covariance estimator)
#'
#' Constructs the V matrix involving inner products of functional predictors
#' (weighted by their Gram matrices) and scalar predictors against response y.
#'
#' @param W A `predictor_hybrid` object.
#' @param y A numeric response vector of length equal to number of samples.
#'
#' @return A matrix representing the corrected cross-covariance structure.
#' @export
get_pre_corrcov <- function(W, y){
  n <- W$n_sample
  K <- W$n_functional

  # building blocks
  ## C_J : C multiplied by J. Scalable to arbitrary K
  predictor_first <- W$functional_list[[1]]
  C_J <- t(predictor_first$coefs) %*% W$gram_list[[1]]
  for (i in 2:K){
    predictor_now <- W$functional_list[[i]]
    C_J <- cbind(
      C_J,
      t(predictor_now$coefs) %*% W$gram_list[[i]]
    )
    }

@@

@@

#

>>>>

### subset_predictor_hybrid

```{r}
#' Extract a single observation from a predictor_hybrid object
#'
#' @param W A predictor_hybrid object with multiple samples.
#' @param i Integer index of the sample to extract.
#' @return A single-sample predictor_hybrid object.
subset_predictor_hybrid <- function(W, i) {
  new_Z <- matrix(W$Z[i, ], nrow = 1)
  new_functional_list <- lapply(W$functional_list, function(fdobj) {
    fd(coef = matrix(coef(fdobj)[, i], ncol = 1), basisobj = fdobj$basis)
  })
  predictor_hybrid(Z = new_Z, functional_list = new_functional_list)
}

@@

###
```{r}
get_CJ <- function(W){
  n <- W$n_sample
  K <- W$n_functional
  ## C_J : C multiplied by J. Scalable to arbitrary K
  predictor_first <- W$functional_list[[1]]
  C_J <- t(predictor_first$coefs) %*% W$gram_list[[1]]
  for (i in 2:K){
    predictor_now <- W$functional_list[[i]]
    C_J <- cbind(
      C_J,
      t(predictor_now$coefs) %*% W$gram_list[[i]]
    )
    }
    return(C_J)
}
```

### get_pre_corrcov
construct the matrix V_star:
$$
    V^\ast = n^{-2}\begin{bmatrix}
         J \widetilde{C}^\top \mathbf{y}\mathbf{y}^\top \widetilde{C} J & J \widetilde{C}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z} \\
        \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C} J &   \widetilde{Z}^\top \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{Z}
    \end{bmatrix} = n^{-2} J^* \widetilde{C}^{*\top} \widetilde{\mathbf{y}} \widetilde{\mathbf{y}}^\top \widetilde{C}^{*} J^*
    \in \mathbb{R}^{(MK+p) \times (MK+p)},
$$
where $J$ is defined in `get_gram_matrix_block`
equation (7), 
```{r}
#' Compute V matrix (pre-corrected covariance estimator)
#'
#' Constructs the V matrix involving inner products of functional predictors
#' (weighted by their Gram matrices) and scalar predictors against response y.
#'
#' @param W A `predictor_hybrid` object.
#' @param y A numeric response vector of length equal to number of samples.
#'
#' @return A matrix representing the corrected cross-covariance structure.
#' @export
get_pre_corrcov <- function(W, y){
  n <- W$n_sample
  K <- W$n_functional

  # building blocks
  ## C_J : C multiplied by J. Scalable to arbitrary K
  C_J <- get_CJ(W)

@@

```{r}
get_xi_direct <- function(W, y) {
  K <- W$n_functional
  C_J <- get_CJ(W)
  A <- cbind(C_J, W$Z)                      # shape: n × d
  b <- t(y) %*% A / W$n_sample   # equivalent to V_star

  J_star <- get_gram_matrix_block(W)
  alpha <- solve(J_star, t(b))                # unconstrained maximizer
  xi_hat <- alpha / sqrt(t(alpha) %*% J_star %*% alpha)  # normalize

  return(xi_hat)
}
```

```{r}
  C_J <- get_CJ(W)
  A <- cbind(C_J, W$Z)                      # shape: n × d
  b <- t(y) %*% A / W$n_sample   # equivalent to V_star

  J_star <- get_gram_matrix_block(W)
  alpha <- solve(J_star, t(b))                # unconstrained maximizer
  xi_hat <- alpha / sqrt(sum((alpha) * (J_star %*% alpha) )) # normalize
t(y) %*% A %*% xi_hat
```

```{r}
 V_star <- get_pre_corrcov(W, y)
L <- chol(J_star)
  invL <- Matrix::chol2inv(L)
  A <- V_star %*% t(invL)       # A = V* × t(inv(L))
  E <- invL %*% A               # E = inv(L) × A

  xi_star <- t(invL) %*% e      # xi_star solves t(L) xi = e
```

@@

### get_pls_comp_simple
```{r}
#' Compute First PLS Component (Direct Hilbert Averaging)
#'
#' Computes the first PLS component as a normalized weighted average of predictor objects,
#' where weights are given by the centered response values.
#'
#' @param W A `predictor_hybrid` object containing `n_sample` observations.
#' @param y A numeric response vector of length `n_sample`.
#'
#' @return A normalized `predictor_hybrid` object representing the first PLS component.
#' @export
get_pls_comp_simple <- function(W, y) {
  n <- W$n_sample
  if (length(y) != n) stop("Length of y must match number of samples in W.")

  # Center the response (optional, but often improves interpretability)
  y_centered <- y - mean(y)

  # Initialize zero element
  xi_sum <- scalar_mul.predictor_hybrid(subset_predictor_hybrid(W, 1), y_centered[1])

  # Accumulate weighted sum: sum_i y_i * W_i
  for (i in 2:n) {
    Wi <- subset_predictor_hybrid(W, i)
    xi_sum <- add.predictor_hybrid(xi_sum, Wi, alpha = y_centered[i])
  }

  # Normalize
  xi_sum <- scalar_mul.predictor_hybrid(xi_sum, 1/n)
  norm_val <- sqrt(inprod.predictor_hybrid(xi_sum))
  xi_hat <- scalar_mul.predictor_hybrid(xi_sum, 1 / norm_val)

  return(xi_hat)
}

```{r}
library(fda)

set.seed(011111)
n_sample <- 100
n_scalar <- 3

# --- Step 1: Zero functional predictors ---
zero_basis <- create.bspline.basis(c(0, 1), nbasis = 4)
zero_fd <- fd(coef = matrix(0, nrow = 4, ncol = n_sample), basisobj = zero_basis)
functional_list <- list(zero_fd, zero_fd)  # two zero-valued functional predictors

# --- Step 2: Simulate scalar predictors and response ---
Z <- matrix(rnorm(n_sample * n_scalar), ncol = n_scalar)
Z <- scale(Z, center = TRUE, scale = FALSE)
W <- predictor_hybrid(Z = Z, functional_list = functional_list)

beta_true <- rnorm(n_scalar)
y <- Z %*% beta_true + rnorm(n_sample, sd = 0.1)
y <- scale(y, center = TRUE, scale = FALSE)

# --- Step 3: Compute constraint matrix ---
lambda <- rep(0, W$n_functional)
L <- chol(get_constraint_matrix_hybrid(W, lambda))

# --- Step 4: PLS with get_pls_comp ---
pls <- get_pls_comp(W, y, L)
xi_hat <- pls$xi_hat

# --- Step 5: PLS via raw Euclidean eigenproblem ---
V_star_vec <- as.vector(t(Z) %*% y)     # equivalent to colSums(Z * y)

J <- crossprod(Z)                       # J = Z^T Z
L_Z <- chol(J)
invL <- solve(L_Z)
E <- invL %*% tcrossprod(V_star_vec) %*% t(invL)

eig <- eigen(E)
e <- eig$vectors[, 1]
xi_manual <- backsolve(L_Z, e, transpose = TRUE)  # xi = L^T \ e

# --- Step 6: get_pls_comp_simple ---
xi_simple <- get_pls_comp_simple(W, y)
xi_simple_vec <- tail(xi_simple$Z, 1)
xi_direct <- get_xi_direct(W, y)
# --- Step 7: Compare all three scalar parts ---

# From get_pls_comp
xi_hat_vec <- tail(xi_hat$Z, 1)

# Cosine similarities
cos_eig_manual <- sum(xi_manual * xi_hat_vec) / (sqrt(sum(xi_manual^2)) * sqrt(sum(xi_hat_vec^2)))
cos_simple_manual <- sum(xi_manual * xi_simple_vec) / (sqrt(sum(xi_manual^2)) * sqrt(sum(xi_simple_vec^2)))
cos_eig_simple <- sum(xi_simple_vec * xi_hat_vec) / (sqrt(sum(xi_simple_vec^2)) * sqrt(sum(xi_hat_vec^2)))

# --- Step 8: Report ---
cat("Cosine similarity (get_pls_comp vs manual eig):", cos_eig_manual, "\n")
cat("Cosine similarity (get_pls_comp_simple vs manual eig):", cos_simple_manual, "\n")
cat("Cosine similarity (get_pls_comp vs get_pls_comp_simple):", cos_eig_simple, "\n")

```{r}
library(fda)

# --- Step 1: Simulate Data ---
set.seed(777)

n_sample <- 100
n_scalar <- 2
n_basis <- 5
n_functional <- 2

# Scalar predictors Z (centered)
Z <- matrix(rnorm(n_sample * n_scalar), ncol = n_scalar)
Z <- scale(Z, center = TRUE, scale = FALSE)

# Create functional predictors (centered)
basis <- create.bspline.basis(c(0, 1), nbasis = n_basis)
functional_list <- list()
for (j in 1:n_functional) {
  coefs <- matrix(rnorm(n_basis * n_sample), nrow = n_basis)
  fdobj <- fd(coef = coefs, basisobj = basis)
  mean_fd <- mean.fd(fdobj)
  mean_coef <- coef(mean_fd)
  centered_coefs <- coefs - matrix(mean_coef, nrow = n_basis, ncol = n_sample)
  functional_list[[j]] <- fd(coef = centered_coefs, basisobj = basis)
}

# --- Step 2: Construct hybrid object ---
W <- predictor_hybrid(Z = Z, functional_list = functional_list)

# --- Step 3: Simulate and center response ---
total_coef_len <- sum(W$n_basis_list) + W$n_scalar
true_coef <- rnorm(total_coef_len)

# Flatten design matrix
Xmat <- matrix(NA, n_sample, total_coef_len)
for (i in 1:n_sample) {
  W_i <- subset_predictor_hybrid(W, i)
  func_coefs <- do.call(c, lapply(W_i$functional_list, function(fdobj) as.vector(coef(fdobj))))
  Xmat[i, ] <- c(func_coefs, W_i$Z)
}

y <- Xmat %*% true_coef + rnorm(n_sample, sd = 0.1)
y <- as.vector(scale(y, center = TRUE, scale = FALSE))

# --- Step 4: Compute constraint matrix ---
lambda <- rep(0, W$n_functional)
constr_mat <- get_constraint_matrix_hybrid(W, lambda)
L <- chol(constr_mat)

# --- Step 5: Run eigen-based PLS ---
W_template <- W  # prevent mutation
pls_eig_result <- get_pls_comp(W, y, L)
xi_hat_eig <- pls_eig_result$xi_hat

# --- Step 6 (Revised): Manually compute direct PLS component and wrap with predictor_hybrid_from_coef ---

# Scalar part

# Functional part: stack the weighted coefficients
func_weighted_vec <- unlist(lapply(functional_list, function(fd_j) {
  coef(fd_j) %*% y  # nbasis × 1
}))

# Full coefficient vector
Z_weighted <- colSums(Z * y)

xi_raw_vec <- c(func_weighted_vec, Z_weighted)

# Wrap into predictor_hybrid object using format template
xi_hat_direct <- predictor_hybrid_from_coef(format = W_template, coef = xi_raw_vec)

# Normalize
norm_val <- sqrt(inprod.predictor_hybrid(xi_hat_direct))
xi_hat_direct <- scalar_mul.predictor_hybrid(xi_hat_direct,  1 / norm_val)

xi_new <- get_xi_direct(W, y)

# --- Step 7: Compare and evaluate ---
ip <- inprod.predictor_hybrid(xi_hat_eig, xi_hat_direct)
norm1 <- sqrt(inprod.predictor_hybrid(xi_hat_eig))
norm2 <- sqrt(inprod.predictor_hybrid(xi_hat_direct))
cos_theta <- ip / (norm1 * norm2)

cat("Cosine similarity between eigen-based and direct PLS components:", cos_theta, "\n")
cat("Projection score (eig):", sum(inprod.predictor_hybrid(xi_hat_eig, W) * y), "\n")
cat("Projection score (direct):", sum(inprod.predictor_hybrid(xi_hat_direct, W) * y), "\n")

# --- Step 8: Plot functional coefficients ---
par(mfrow = c(n_functional, 1), mar = c(4, 4, 2, 2))
for (j in 1:n_functional) {
  plot(xi_hat_eig$functional_list[[j]], col = "blue", ylim = c(-1, 1),
       main = paste("Functional Component", j), ylab = "Coefficient", xlab = "t")
  lines(xi_hat_direct$functional_list[[j]], col = "red", lty = 2)
  legend("topright", legend = c("Eigen-based", "Direct"), col = c("blue", "red"), lty = 1:2)
}

## Residualization

### LSE_hybrid
```{r}
LSE_hybrid <- function(W, rho, tau) {
  C_star_J_star <- matrix(NA, nrow = W$n_sample, ncol = sum(W$n_basis_list) + W$n_scalar)
  for (ii in (1 : W$n_functional)){
    cumsum_n_basis_now <- sum(W$n_basis_list[1:ii])
    n_basis_now <- W$n_basis_list[ii]
    C_star_J_star[, (cumsum_n_basis_now - n_basis_now + 1) :cumsum_n_basis_now] <-
      t(W$functional_list[[ii]]$coefs) %*% W$jacobian_list[[ii]]
  }
  C_star_J_star[ ,(sum(W$n_basis_list) + 1) : ncol(C_star_J_star)] <- W$Z
  rho_t_C_star_J_star <- t(rho) %*% C_star_J_star

  # matrices in the statement of Proposition 4
  J_star <-as.matrix(get_jacobian_hybrid(W))
  d_hat_star <- t( #basis coefficient of hybrid regression coefficient
    solve(
      t((norm(rho, type="2"))^2 * J_star),
      t(rho_t_C_star_J_star)
      )
    )
  delta_hat <- hybrid_from_coef(W, d_hat_star) #hybrid regression coefficient
  return(delta_hat)
  }
```

### nipals_pen_hybrid

- depends on:
  - 'get_constraint_matrix_hybrid'
  - 'get_pls_comp'
  - 'hybrid_inner_prod'
  - 'LSE_hybrid'
